### 2022-09-12 ms/manuscript.Rmd
make template

---
title: "manuscript"
author: "Daniel Suh"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

# Introduction

# Methods

# Results

# Discussion

# References