### 2023-01-31 scripts/workshop/sem.Rmd
add more notes

---
title: "sem"
output: html_document
date: "2022-12-06"
---

```{r, message = F}
library(here)

library(tidyverse)
library(magrittr)
library(piecewiseSEM)
library(nlme)
library(lmerTest)
library(mgcv)
library(dagitty)

comm_data <- readRDS(here("processed_data","comm_data.rds"))
```

Make lagged data - I'll use this later to compare results using lagged and not lagged data
```{r}
prev_cc <- comm_data

order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]

#this removes months out of sequence
order <- order[-c(4,14,47,90),]

order %<>% add_column(lag_cc = NA, lag_size = NA, lag_temp = NA)

#create new column that includes previous month's value for cc
for(n in 2:nrow(order)){
  order$lag_cc[n] <- order$cc[n-1]
  order$lag_size[n] <- order$size[n-1]
  order$lag_temp[n] <- order$MeanWaterTempPredC[n-1]
}

#remove the first entry for each wetland to remove the carryover from the last wetland
clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)

```

This is just a way to visualize the SEM

```{r}
prev_mod <- dagitty("dag{
  temp -> cc
  temp -> size
  temp -> prev
  cc -> prev
  size -> prev
             }")

plot(graphLayout(prev_mod))

```

Just some histograms to check the distributions of the data

```{r}
comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC)) + geom_histogram()
comm_data %>% ggplot(.,aes(x=size)) + geom_histogram()
comm_data %>% ggplot(.,aes(x=log10(size))) + geom_histogram()
comm_data %>% ggplot(.,aes(x=cc)) + geom_histogram()
comm_data %>% ggplot(.,aes(x=log(cc))) + geom_histogram()
comm_data %>% ggplot(.,aes(x=Prevalence)) + geom_histogram()
comm_data %>% ggplot(.,aes(x=log10(Prevalence+1))) + geom_histogram()

```

And some basic plots

```{r}
comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = cc)) + geom_point() + geom_smooth()
comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = log10(size))) + geom_point() + geom_smooth()
comm_data %>% ggplot(.,aes(x=MeanWaterTempPredC, y = Prevalence)) + geom_point() + geom_smooth()
comm_data %>% ggplot(.,aes(x=log10(cc), y = Prevalence)) + geom_point() + geom_smooth(method = "lm")
comm_data %>% ggplot(.,aes(x=log10(size), y = Prevalence)) + geom_point() + geom_smooth(method = "lm")
```

Covariance-based SEM
```{r}

```

How to model nonlinear effect of temperature?
- hump-shaped relationship may be expected with temperature

cc ~ mean water temp; size ~ mean water temp; prevalence ~ cc + size + mean water temp
```{r}
cc_temp <- lm(log10(cc) ~ (MeanWaterTempPredC), data = comm_data)
size_temp <- lm(log10(size) ~ (MeanWaterTempPredC), data = comm_data)
prev_cc <- lm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
prev_temp_cc <- glm(Prevalence ~ log10(cc) + log10(size) + MeanWaterTempPredC, data = comm_data)

sem <- psem(cc_temp, size_temp, prev_cc)
sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
```

```{r}
comm_data %<>% filter(MeanWaterTempPredC>0)

cc_temp <- gam(log10(cc) ~ s(MeanWaterTempPredC), data = comm_data)
size_temp <- lm(log10(size) ~ (MeanWaterTempPredC), data = comm_data)
prev_cc <- lm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
prev_temp_cc <- glm(Prevalence ~ log10(cc) + log10(size) + MeanWaterTempPredC, data = comm_data)

summary(cc_temp)

#fix error
#sem <- psem(cc_temp, size_temp, prev_cc)
#sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
```

Using lagged data
```{r}
lagged_cc_temp <- lm(lag_cc ~ lag_temp, data = clean)
lagged_size_temp <- lm(log10(lag_size) ~ lag_temp + lag_cc, data = clean)
lagged_prev_cc <- lm(Prevalence ~ lag_cc + log10(lag_size), data = clean)
lagged_prev_temp_cc <- lm(Prevalence ~ lag_cc + log10(lag_size) + lag_temp, data = clean)

sem_lagged <- psem(lagged_cc_temp, lagged_size_temp, lagged_prev_cc)
sem_1_lagged <- psem(lagged_cc_temp, lagged_size_temp, lagged_prev_temp_cc)
```

```{r}
basisSet(sem)
dSep(sem, conditioning = T)
fisherC(sem)
coefs(sem, intercepts = T)
rsquared(sem)
```

```{r}
basisSet(sem_1)
dSep(sem_1, conditioning = T)
fisherC(sem_1)
coefs(sem_1, intercepts = T)
rsquared(sem_1)
```

```{r}
summary(sem)
plot(sem)
plot(sem, return = T)
```

```{r}
summary(sem_1)
plot(sem_1)
plot(sem_1, return = T)
```

```{r}
summary(sem_lagged)
plot(sem_lagged)
plot(sem_lagged, return = T)
```

```{r}
summary(sem_1_lagged)
plot(sem_1_lagged)
plot(sem_1_lagged, return = T)
```

Lefcheck example
This is example data and code that I got from Lefcheck's tutorial
https://cran.r-project.org/web/packages/piecewiseSEM/vignettes/piecewiseSEM.html#worked-example

```{r}
data(shipley)
```

```{r}
dd_lat <-  lme(DD~lat, random = ~1|site/tree, na.action = na.omit, 
    data = shipley)

date_dd <-lme(Date~DD, random = ~1|site/tree, na.action = na.omit, 
    data = shipley)

growth_date <-  lme(Growth~Date, random = ~1|site/tree, na.action = na.omit, 
    data = shipley)

live_growth <-  glmer(Live~Growth+(1|site)+(1|tree), 
    family=binomial(link = "logit"), data = shipley)
```

```{r}
shipley_sem <- psem(dd_lat, date_dd, growth_date, live_growth)
```

```{r}
summary(shipley_sem)
```