### 2023-03-29 area_UC.Rmd
Create area_UC.Rmd

---
title: "Area Under Curve"
author: "Daniel Suh"
date: "`r Sys.Date()`"
output: html_document
---

```{r, message = F}
library(here)

source(here("base","src.R"))

library(pracma)
```

Look at area under curve for all variables for each community (i.e. site-month)
```{r}
comm_data <- readRDS(here("processed_data","comm_data.rds"))
```

```{r}
comm_data %<>% mutate(month_n = as.numeric(gsub("Month","",Month.1)),
                      WetAltID = as.factor(WetAltID))
areas <- comm_data %>% group_by(WetAltID) %>% summarize(abund = trapz(month_n, size),
                                                        temp = trapz(month_n, MeanWaterTempPredC),
                                                        cc = trapz(month_n, cc),
                                                        prev = trapz(month_n, Prevalence))
```

```{r}
areas %>% ggplot(.,aes(x=abund,y=prev)) + geom_point() + geom_smooth(method = "lm")
areas %>% ggplot(.,aes(x=cc,y=prev)) + geom_point() + geom_smooth(method = "lm")
areas %>% ggplot(.,aes(x=temp,y=prev)) + geom_point() + geom_smooth(method = "lm")
```

```{r}
abund_lm <- lm(prev ~ log(abund), areas)
cc_lm <- lm(prev ~ cc, areas)
temp_lm <- lm(prev ~ temp, areas)
```